---
title: "Project - Group 10"
author: "Anthony Desdha, Santiago Fonnegra, Tim Haynes, Hayden Post, Vyvy Tran, Valeria Valle, Katie Yang"
date: "SDS 322E"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r message=FALSE}

library(tidyverse)

# select data Water_Quality <- read.csv("Water_Quality.csv")

Water_Quality <- read.csv("~/Project 10/Water_Quality.csv", header=FALSE)

Water_Quality_new <- Water_Quality %>% separate(V4, into = c("Date", "Time", "Meridium"), sep = " ")

Water_Quality_new <- Water_Quality_new %>% separate(Date, into = c("Month", "Day", "Year"), sep = "/") 

```

## Description

**Analyzing the relationship between pathogenic bacteria and organic nutrients present in freshwater samples taken in Austin.**

## Goals

**Our goals for this project is to observe how different organic nutrients and different locations can affect E.coli presence. Additionally, we also want to be able to analyze a data set and create/conclude data visualizations and answer our hypothesis.**

## Link to Data

**https://drive.google.com/file/d/13872S6iGB2MThwLKrlwKPZ2B5RK_kt5j/view?usp=sharing**

## Information on the Data

## Data Visualizations

```{r}
library(ggplot2)

#Visualization 1

Water_Quality_new %>% group_by(V17)%>% filter(V17 == "E COLI BACTERIA") %>% ggplot(aes(x=V6)) + geom_bar(alpha=1, size=2)+ theme(axis.text.x = element_text(angle = 45, vjust = 0.7))

# Visualization 2
```

## Hypothesis

**1:**
**2:**

## References

**This data originally came from a larger data set describing the same variables. This originally data set is called "Water Quality Sampling Data" located on the data.austintexas.gov website.**



When you click the **Knit** button a document will be generated that includes both content as well as the output of any embedded R code chunks within the document. You can embed an R code chunk like this:

```{r cars}
summary(cars)
```

## Including Plots

You can also embed plots, for example:

```{r pressure, echo=FALSE}
plot(pressure)
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
