---
title: "Project - Group 10"
author: "Anthony Desdha, Santiago Fonnegra, Tim Haynes, Hayden Post, Vyvy Tran, Valeria Valle, Katie Yang"
date: "SDS 322E"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Description

Analyzing the relationship between pathogenic bacteria and organic nutrients present in freshwater samples taken in Austin. 

# select data Water_Quality <- read.csv("Water_Quality.csv")
# fic data from above to reflect the subset data we pulled

## Goals

## Link to Data
```{r cars}
Water_Quality_new <- Water_Quality %>% separate(SAMPLE_DATE, into = c("Date", "Time", "Meridium"), sep = " ") Water_Quality_new <- Water_Quality_1 %>% separate(Date, into = c("Month", "Day", "Year"), sep = "/") #Visualization 1 Water_Quality_new %>% group_by(PARAMETER)%>% filter(PARAMETER == "E COLI BACTERIA") %>% ggplot(aes(x=WATERSHED)) + geom_bar(alpha=1, size=2)+ theme(axis.text.x = element_text(angle = 45, vjust = 0.7))
```

## Information on the Data

## Data Visualizations

## Hypothesis

## References

When you click the **Knit** button a document will be generated that includes both content as well as the output of any embedded R code chunks within the document. You can embed an R code chunk like this:

```{r cars}
summary(cars)
```

## Including Plots

You can also embed plots, for example:

```{r pressure, echo=FALSE}
plot(pressure)
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
